---
title: "penguins_linear_regression.qmd"
format: html
editor: visual
---

# Challenge: Modeling Penguin Flipper Length with Bayesian Regression

You're going to build a Bayesian Normal regression model predicting flipper length (`flipper_length_mm`) from bill length (`bill_length_mm`), using the `penguins_bayes` dataset. Work through each section in order, and answer every numbered question directly in your qmd, right under the code that produces it.

## 1. Load the data

```{r}
library(tidyverse)
library(bayesrules)
library(rstanarm)
library(plotly)
library(report)

data <- bayesrules::penguins_bayes
```

## 2. Look before you model

```{r}
view(data)
```

```{r}
graph_1 <- ggplot(data = data, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram() +
  theme_bw()

ggplotly(graph_1)
```

### What does the histogram say?

- The graph shows the different species and how many has the same flipper length.

```{r}
graph_2 <- ggplot(data = data, aes(x = bill_length_mm, y = flipper_length_mm)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplotly(graph_2)
```

### What does the scatterplot say?

- This graph shows that both x and y values are increasing and how strong their connects are.

## 3. Translate these three beliefs into priors

Read each statement carefully — you need to convert plain language into specific numbers.

> **Typical flipper length.** Across all three species in this dataset, a reasonable guess for average flipper length is around 200mm — with a plausible range of roughly 180mm to 220mm.

> **The bill-to-flipper relationship.** Bill length and flipper length are both rough proxies for body size, so for every 1mm increase in bill length, flipper length typically increases by around 3.5mm — with a plausible range of roughly 1.5mm to 5.5mm.

> **Leftover variability.** Even penguins with identical bill lengths won't have identical flippers. A reasonable guess for the typical leftover spread is around 12.5mm.

**How to convert each one:**

- For a Normal prior, "around X, plausibly between A and B" means: X is your mean, and the width from A to B is your **95% range**, which equals **mean ± 1.96 × SD**. Solve for SD.
- For sigma's Exponential prior, there's no range to convert — just a single target mean. Recall **E(sigma) = 1/rate**, so `rate = 1/E(sigma)`.

Write your five numbers as plain variables before moving on:

```{r}
m0 <- 200   # centered intercept mean
s0 <- 10.2  # centered intercept SD
m1 <- 3.5   # slope mean
s1 <- 1     # slope SD
l  <- 0.08  # sigma's exponential rate
```

## 4. Fit the model

```{r}

model <- stan_glm(
  flipper_length_mm ~ bill_length_mm, data = data,
  family = gaussian,
  prior_intercept = normal(m0, s0),
  prior = normal(m1, s1),
  prior_aux = exponential(l),
  chains = 4, iter = 2000, seed = 84735
)

```

```{r}
model

summary(model)

plot_normal(1.7, 0.051)

report::report(model)
```

```{r}

posterior <- as.data.frame(model)
beta0c_draws <- posterior$`(Intercept)` + posterior$bill_length_mm * mean(data$bill_length_mm)
```

<https://www.mdpi.com/2072-4292/13/14/2677>

They compared their bleaching prediction ability using spatiotemporal Bayesian hierarchical models (analyze data that vary across both space and time, while accounting for uncertainty using probability) and sensitivity–specificity analyses.

The `echo: false` option disables the printing of code (only output is displayed).
